Supplement: Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-09 supplement.tex
tiny structure change for text examples
edits re. ELBO

If we were to maximize the ELBO bound to perform inference in
our generative model,
then, during DSL induction, we would seek a new $(\mathcal{D}^*,\theta^*)$ maximizing the following lower bound on the likelihood (along an unimportant regularizing term on the DSL):
\begin{align}
  &  \sum_{x\in X}\expect_{p\sim Q_x}\left[\log \probability[p|\mathcal{D}^*,\theta^*] \right] \label{EM}\\
  & Q_x(p)\triangleq \probability[p|x,\mathcal{D},\theta]
\end{align}
where $(\mathcal{D},\theta)$ is our current estimate of the generative
model.  These equations fall out of an EM-style derivation, and one
could replace $Q_x(p)$ with the recognition model $q(p|x)$, either
using importance sampling (so the expectation in Eq.~\ref{EM} is taken
over $q$ and we reweigh using $Q_x$) or by directly using $q$ as our approximate posterior over the program that solves task $x$.

How the recognition model is worked into this is unimportant to the reason why we do not maximize a bound of this form:
this bound takes an expectation over the \emph{previous} sleep cycle's
posterior over programs,
and so the approximate posterior $Q_x$ at the next iteration
ends up being very close to previous approximate posterior.
Intuitively,
we want the DSL induction to be a function \emph{only} of the programs that we have found,
and \emph{not} be a function of how the previous generative model weighed them.
In practice,
we found that maximizing EM-style bounds, like the ELBO,
leads to a kind of hysteresis effect,
where the next generative model too closely matches the previous one,
causing the algorithm to easily become trapped in local optima.

@@

We automatically generated 600 string editing problems split 75/25 train/test.
The following are representative problems:
\pagebreak

>>>>

If we were to maximize the ELBO bound to perform inference in
our generative model,
then, during DSL induction, we would seek a new $(\mathcal{D}^*,\theta^*)$ maximizing the following lower bound on the likelihood (along with an unimportant regularizing term on the DSL):
\begin{align}
  &  \sum_{x\in X}\expect_{p\sim Q_x}\left[\log \probability[p|\mathcal{D}^*,\theta^*] \right] \label{EM}\\
  & Q_x(p)\triangleq \probability[p|x,\mathcal{D},\theta]
\end{align}
where $(\mathcal{D},\theta)$ is our current estimate of the generative
model.  These equations fall out of an EM-style derivation, and one
could replace $Q_x(p)$ with the recognition model $q(p|x)$, either
using importance sampling (so the expectation in Eq.~\ref{EM} is taken
over $q$ and we reweigh using $Q_x$) or by directly using $q$ as our approximate posterior over the program that solves task $x$.

We do not maximize a bound of this form because it
takes an expectation over the \emph{previous} sleep cycle's
posterior over programs,
so the approximate posterior $Q_x$ at the next iteration
ends up being very close to previous approximate posterior.
Intuitively,
we want the DSL induction to be a function \emph{only} of the programs that we have found,
and \emph{not} be a function of how the previous generative model weighed them.
In practice,
we found that maximizing EM-style bounds, like the ELBO,
leads to a kind of hysteresis effect,
where the next generative model too closely matches the previous one,
causing the algorithm to easily become trapped in local optima.

@@

We automatically generated 600 string editing problems split 75/25 train/test.
The following are representative problems:
